---
title: "Heatmap figure"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## Packages

```{r}
library(magrittr)
library(stringr)
library(purrr)
library(tidyr)
library(corrplot)
library(dplyr)
```

## Data

Reading the data:

```{r}
data <- readxl::read_excel("Data_EOs.xlsx")
```

Reformating the names of the products:

```{r}
names(data) <- data %>%
  names() %>%
  str_replace("Activo", "product A") %>% 
  str_replace("Cinamon", "Cinnamon") %>% 
  str_replace(" ", "_")
```

## The functions we need

This function makes a formula from `x` and `y`, correcting for pathogens:

```{r}
make_formula <- function(y, x) {
  formula(paste0("log2(", y, ") ~  Pathogen + log2(", x, ")"))
}
```

This function returns a value of correlation coefficient (corrected for pathogens)
together with p values from the LRT:

```{r}
cor_coef <- function(x, y) {
  if(x == y) return(c(corcoef = 1, p_value = 0, coefval = 1))
  mod <- glm(make_formula(x, y), data = data)
  ano <- anova(mod, test = "LRT")
  c(corcoef = sign(unname(last(coef(mod)))) * last(ano$Deviance) / first(tail(ano$`Resid. Dev`, 2)),
    p_value = last(ano$`Pr(>Chi)`),
    coefval = last(coef(mod)))
}
```

Let's try them:

```{r}
cor_coef("COL", "ENR")
```

and:

```{r}
cor_coef("COL", "COL")
```

## Let's run for all the combinations

Let's extract the names of the essential oils and antibiotics:

```{r}
ab_eo <- rev(setdiff(names(data), c("No.", "Pathogen")))
```

Another set of names where `_` is replaced by a space:

```{r}
ab_eo2 <- str_replace(ab_eo, "_", " ")
```

Let's make all the combinations of essential oils and antibiotics:

```{r}
combinations <- expand_grid(v1 = ab_eo, v2 = ab_eo)
```

Calculating the correlation coefficients and p values for all the combinations
of essential oils and antibiotics:

```{r}
corval <- combinations %>% 
  mutate(a = map2(v1, v2, cor_coef)) %>% 
  unnest_wider(a)
```

Which gives:

```{r}
corval
```

This function arranges the results into a named matrix for later use by the
`corrplot::corrplot()` function:

```{r}
put_in_shape <- function(x) {
  x %>%
    select(-v1) %>% 
    as.matrix() %>% 
    `rownames<-`(ab_eo2) %>% 
    `colnames<-`(ab_eo2)
}
```

Using this function to generate a matrix of p values:

```{r}
pmat <- corval %>% 
  select(-corcoef, -coefval) %>% 
  pivot_wider(names_from = v2, values_from = p_value) %>% 
  put_in_shape()
```

a matrix of correlation coefficients:

```{r}
cormat <- corval %>% 
  select(-p_value, -coefval) %>% 
  pivot_wider(names_from = v2, values_from = corcoef) %>% 
  put_in_shape()
```

and a matrix of regression coefficients:

```{r}
coefmat <- corval %>% 
  select(-p_value, -corcoef) %>% 
  pivot_wider(names_from = v2, values_from = coefval) %>% 
  put_in_shape()
```

## Correlation matrix figure

Replacing non-significant p values by `NA`:

```{r}
coefmat[which(pmat > .05)] <- NA
```

Generating a color palette:

```{r}
col2 <- colorRampPalette(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
                           "#FDDBC7", "#FFFFFF", "#D1E5F0", "#92C5DE",
                           "#4393C3", "#2166AC", "#053061"))
```

Making the plot, in 2 steps:

```{r fig.height=11, fig.width=11, error = TRUE}
corrplot(cormat, "upper", diag = FALSE, p.mat = pmat, method = "color",
         sig.level = 1, cl.pos = "b", tl.col = "black", col = col2(20))
corrplot(coefmat, "upper", diag = FALSE, method = "number", bg = "transparent",
         cl.pos = "n", col = "black", number.font = 1, na.label = "",
         is.corr = FALSE, add = TRUE)
```
